---
title: "Bayesian Regression Analysis of Skewed Tensor Responses"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bayesian Regression Analysis of Skewed Tensor Responses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
options(knitr.table.format = 'markdown')
```

Inkoo Lee

Last Updated - April 22, 2022

## Paper 

- Lee, I., Sinha, D., Mai, Q., Zhang, X. & Bandyopadhyay, D. (2021+). *Bayesian Regression Analysis of Skewed Tensor Responses.* [[Github]](https://github.com/InkooLee/BSTN)


## Skewed tensor response model
For an order-$K$ tensor response $\mathcal{Y}_i \in \mathbb{R}^{d_1 \times \cdots \times d_K}$ with a vector of covariates $\textbf{x}_i \in \mathbb{R}^p$, we consider a tensor response regression model 
\begin{equation}\label{eq1}
    \mathcal{Y}_{i} = \mathcal{B}\bar{\times}_{(K+1)} \textbf{x}_{i} + \mathcal{E}_{i},~~for~~ i=1,\ldots,n,~ 
\end{equation}
where $\mathcal{B} \in \mathbb{R}^{d_1 \times \cdots \times d_K \times p}$ is an order-$(K+1)$ unknown tensor of regression coefficients, $\bar{\times}_{(K+1)}$ is the $(K+1)$-mode vector product, and the error $\mathcal{E}_{i} \in \mathbb{R}^{d_1 \times \cdots \times d_K}$ is an order-$K$ tensor. We model the skewness in 
the distribution of $\mathcal{Y}$ via 
\begin{equation}\label{eq2}
 \mathcal{E}_{i} = |\mathcal{Z}_{2i}|\times_{K} \pmb{\Lambda} + \mathcal{Z}_{1i}, 
\end{equation}
where $\pmb{\Lambda} = diag(\lambda_1,\ldots,\lambda_{d_K}) \in \mathbb{R}^{d_K \times d_K}$ with skewness parameters $\pmb{\lambda}=(\lambda_1,\ldots,\lambda_{d_K})$ and $|\mathbf{M}|$ denotes the
matrix whose each element of is the absolute value of the corresponding element of matrix $\mathbf{M}$. 

## Tensor Spike-and-Slab prior
\begin{equation}\label{eq:IdentifyingCoef}
\begin{aligned}
&\mathcal{B}_{i_1 i_2 i_3 j} = \eta_{i_3j} + \gamma_{i_1i_2i_3 j},~~ \eta_{i_3j} \sim N(0, \tau^2)~~ for~~ \tau > 0
\end{aligned}
\end{equation}
and introduce sparsity via spike-and-slab priors \citep{ishwaran2005spike} as 
\begin{equation}\label{eq:SparsePrior}
\gamma_{i_1i_2i_3j} \overset{iid}{\sim}  (1 - \omega_{i_1
i_2 i_3 j})\delta_0(\gamma_{i_1i_2i_3 j}) +  \omega_{i_1i_2i_3 j}N(0, \nu^2)~~ for~~ \nu > 0,
\end{equation}
where, $\delta_0(\cdot)$ is the indicator function at 0 and $\omega_{i_1i_2i_3 j} \overset{iid}{\sim} Bernoulli(\psi)$ with hyperprior $\psi \sim Beta(a,b)$. For the GAAD study, we choose $a = b = 0.1$, representing a vague hyperprior for $\psi$. 

Since sparsity inducing prior on $\gamma_{i_1i_2i_3j}$ tends to concentrate around zero, $\mathcal{B}$ in \eqref{eq:IdentifyingCoef} is mostly determined by $\eta_{i_3j}$ allowing parsimonious interpretation of covariate effects, while estimates of $\mathcal{B}$ using Tensor Normal prior (with large-scale covariances)  are much more different from each other.

## GAAD data analysis using `BSTN` package


```{r, eval = FALSE}
## Load package
library(BSTN)
```

We fit the BSTN (Bayesian skewed tensor normal) model with tensor SAS (spike-and-slab) prior.

```{r, eval = FALSE}
# n.burn = burn-in period, n.save = number of posterior samples, thin = thinning size
BSTN_SAS <- BSTN_SAS(Y,X,vecy, n.burn = 100, n.save = 1000, thin = 5)
```

### D-Statistic
$D = d_1/|d_2|$ for 
each regression effects $\mathcal{B}_{i_1i_2i_3j}$, where $d_1$ is distance of the Bayesian estimate (posterior median) from null value (0 here) and $d_2$ is the distance of the Bayes estimate from the nearest endpoint of the 95\% CI of the parameter. 
Here, $D > 1.5$ $(D < -1.5)$ implies strong posterior evidence of positive (negative) effect of covariate $x_j$ on the biomarker $i_3$ at site $i_2$ of tooth $i_1$. 
Similarly, the range $-0.8 < D < 0.8$ suggests moderate evidence 
for null hypothesis, and $0.8 \leq  D \leq 1.5$ represents moderate positive effect, and $-1.5 \leq D \leq -0.8$ denotes moderate negative effect of $x_j$ on $\mathcal{Y}_{i_1i_2i_3}$.

The following plot (Figure 1 in the main text) shows D-statistic map of tensor regression coefficients with sparsity inducing prior. 

```{r, echo=FALSE, fig.align='center', fig.width=7, fig.height=3, warning=FALSE}
load("C:/Users/inkku/OneDrive/Desktop/BSTN/GAAD_result/PPD.rda")
load("C:/Users/inkku/OneDrive/Desktop/BSTN/GAAD_result/CAL.rda")

mat<-function(A,j)
{
  Aj<-t(apply(A,j,"c"))
  if(nrow(Aj)!=dim(A)[j])  { Aj<-t(Aj) }
  Aj
}

PPD_B.est <- mat(PPD3$B.est,1)[2:6,] # exclude intercept
CAL_B.est <- mat(CAL$B.est,1)[2:6,] # exclude intercept

PPD2init <- as.matrix(PPD_B.est[1,])
PPD3init <- as.matrix(PPD_B.est[2,])
PPD4init <- as.matrix(PPD_B.est[3,])
PPD5init <- as.matrix(PPD_B.est[4,])
PPD6init <- as.matrix(PPD_B.est[5,])


CAL2init <- as.matrix(CAL_B.est[1,])
CAL3init <- as.matrix(CAL_B.est[2,])
CAL4init <- as.matrix(CAL_B.est[3,])
CAL5init <- as.matrix(CAL_B.est[4,])
CAL6init <- as.matrix(CAL_B.est[5,])

library(latticeExtra)
library(RColorBrewer)
library(lattice)
library(rasterVis)
library(circlize)

t=28;s=6;
#Acol.regions <- c("#FF80FFFF", "#FFFFFF", "#1E90FF")
Acol.regions <- c("#FF80FFFF","#FFB6C1", "#FFFFFF", "#BFFFFFFF", "#1E90FF")

n.iter = 1000
## PPD AGE ##
PPD2 <- matrix(NA,168,n.iter)
for (i in 1:n.iter){
  PPD2[,i] <- PPD2init[(168*(i-1) + 1):(168*i),] 
}

CIPPD2 <- matrix(NA,168,3)

for (i in 1:168){
  CIPPD2[i,] <- quantile(PPD2[i,], c(0.025,0.5,0.975))
}


PPD2_d1 <- as.matrix(CIPPD2[,2]); PPD2_d2 <- matrix(NA, 168, 1)
for (i in 1:168){
  if (PPD2_d1[i,] > 0){
    PPD2_d2[i,] <- PPD2_d1[i,] - CIPPD2[i,1]
  }
  else {PPD2_d2[i,] <- CIPPD2[i,3] - PPD2_d1[i,]}
}
PPD2map <- PPD2_d1/abs(PPD2_d2)
PPD2map <- matrix(PPD2map, t,s)

PPD2map <- levelplot(t(PPD2map),col.regions = Acol.regions,xlab="sites",ylab="teeth", at = c(-4,-1.5,-0.8,0.8,1.5,4),
                     colorkey = FALSE, labels=as.character(c(-4,-1.5,-0.8,0.8,1.5,4)),
                     margin=F, scales = list(draw = FALSE),main = list(label = "PPD_Age", cex = 1))

print(PPD2map, split = c(1, 1, 5, 2), more = TRUE)

## PPD Gender ##
PPD3 <- matrix(NA,168,n.iter)
for (i in 1:n.iter){
  PPD3[,i] <- PPD3init[(168*(i-1) + 1):(168*i),] 
}

CIPPD3 <- matrix(NA,168,3)

for (i in 1:168){
  CIPPD3[i,] <- quantile(PPD3[i,], c(0.025,0.5,0.975))
}


PPD3_d1 <- as.matrix(CIPPD3[,2]); PPD3_d2 <- matrix(NA, 168, 1)
for (i in 1:168){
  if (PPD3_d1[i,] > 0){
    PPD3_d2[i,] <- PPD3_d1[i,] - CIPPD3[i,1]
  }
  else {PPD3_d2[i,] <- CIPPD3[i,3] - PPD3_d1[i,]}
}
PPD3map <- PPD3_d1/abs(PPD3_d2)
PPD3map <- matrix(PPD3map, t,s)

PPD3map <- levelplot(t(PPD3map),col.regions = Acol.regions,xlab="sites",ylab="teeth", at = c(-4,-1.5,-0.8,0.8,1.5,4),
                     colorkey = FALSE, labels=as.character(c(-4,-1.5,-0.8,0.8,1.5,4)),
                     margin=F, scales = list(draw = FALSE),main = list(label = "PPD_Gender", cex = 1))

print(PPD3map, split = c(2, 1, 5, 2), more = TRUE)

## PPD BMI ##
PPD4 <- matrix(NA,168,n.iter)
for (i in 1:n.iter){
  PPD4[,i] <- PPD4init[(168*(i-1) + 1):(168*i),] 
}

CIPPD4 <- matrix(NA,168,3)

for (i in 1:168){
  CIPPD4[i,] <- quantile(PPD4[i,], c(0.025,0.5,0.975))
}


PPD4_d1 <- as.matrix(CIPPD4[,2]); PPD4_d2 <- matrix(NA, 168, 1)
for (i in 1:168){
  if (PPD4_d1[i,] > 0){
    PPD4_d2[i,] <- PPD4_d1[i,] - CIPPD4[i,1]
  }
  else {PPD4_d2[i,] <- CIPPD4[i,3] - PPD4_d1[i,]}
}
PPD4map <- PPD4_d1/abs(PPD4_d2)
PPD4map <- matrix(PPD4map, t,s)

PPD4map <- levelplot(t(PPD4map),col.regions = Acol.regions,xlab="sites",ylab="teeth", at = c(-4,-1.5,-0.8,0.8,1.5,4),
                     colorkey = FALSE, labels=as.character(c(-4,-1.5,-0.8,0.8,1.5,4)),
                     margin=F, scales = list(draw = FALSE),main = list(label = "PPD_BMI", cex = 1))
print(PPD4map, split = c(3, 1, 5, 2), more = TRUE)



## PPD Smoker ##
PPD5 <- matrix(NA,168,n.iter)
for (i in 1:n.iter){
  PPD5[,i] <- PPD5init[(168*(i-1) + 1):(168*i),] 
}

CIPPD5 <- matrix(NA,168,3)

for (i in 1:168){
  CIPPD5[i,] <- quantile(PPD5[i,], c(0.025,0.5,0.975))
}


PPD5_d1 <- as.matrix(CIPPD5[,2]); PPD5_d2 <- matrix(NA, 168, 1)
for (i in 1:168){
  if (PPD5_d1[i,] > 0){
    PPD5_d2[i,] <- PPD5_d1[i,] - CIPPD5[i,1]
  }
  else {PPD5_d2[i,] <- CIPPD5[i,3] - PPD5_d1[i,]}
}
PPD5map <- PPD5_d1/abs(PPD5_d2)
PPD5map <- matrix(PPD5map, t,s)

PPD5map <- levelplot(t(PPD5map),col.regions = Acol.regions,xlab="sites",ylab="teeth", at = c(-4,-1.5,-0.8,0.8,1.5,4),
                     colorkey = FALSE, labels=as.character(c(-4,-1.5,-0.8,0.8,1.5,4)),
                     margin=F, scales = list(draw = FALSE),main = list(label = "PPD_Smoker", cex = 1))
print(PPD5map, split = c(4, 1, 5, 2), more = TRUE)

## PPD HbA1c ##
PPD6 <- matrix(NA,168,n.iter)
for (i in 1:n.iter){
  PPD6[,i] <- PPD6init[(168*(i-1) + 1):(168*i),] 
}

CIPPD6 <- matrix(NA,168,3)

for (i in 1:168){
  CIPPD6[i,] <- quantile(PPD6[i,], c(0.025,0.5,0.975))
}


PPD6_d1 <- as.matrix(CIPPD6[,2]); PPD6_d2 <- matrix(NA, 168, 1)
for (i in 1:168){
  if (PPD6_d1[i,] > 0){
    PPD6_d2[i,] <- PPD6_d1[i,] - CIPPD6[i,1]
  }
  else {PPD6_d2[i,] <- CIPPD6[i,3] - PPD6_d1[i,]}
}
PPD6map <- PPD6_d1/abs(PPD6_d2)
PPD6map <- matrix(PPD6map, t,s)

PPD6map <- levelplot(t(PPD6map),col.regions = Acol.regions,xlab="sites",ylab="teeth", at = c(-4,-1.5,-0.8,0.8,1.5,4),
                     colorkey = FALSE, labels=as.character(c(-4,-1.5,-0.8,0.8,1.5,4)),
                     margin=F, scales = list(draw = FALSE),main = list(label = "PPD_HbA1c", cex = 1))
print(PPD6map, split = c(5, 1, 5, 2), more = TRUE)


## CAL AGE ##
CAL2 <- matrix(NA,168,n.iter)
for (i in 1:n.iter){
  CAL2[,i] <- CAL2init[(168*(i-1) + 1):(168*i),] 
}

CICAL2 <- matrix(NA,168,3)

for (i in 1:168){
  CICAL2[i,] <- quantile(CAL2[i,], c(0.025,0.5,0.975))
}


CAL2_d1 <- as.matrix(CICAL2[,2]); CAL2_d2 <- matrix(NA, 168, 1)
for (i in 1:168){
  if (CAL2_d1[i,] > 0){
    CAL2_d2[i,] <- CAL2_d1[i,] - CICAL2[i,1]
  }
  else {CAL2_d2[i,] <- CICAL2[i,3] - CAL2_d1[i,]}
}
CAL2map <- CAL2_d1/abs(CAL2_d2)
CAL2map <- matrix(CAL2map, t,s)

CAL2map <- levelplot(t(CAL2map),col.regions = Acol.regions,xlab="sites",ylab="teeth", at = c(-4,-1.5,-0.8,0.8,1.5,4),
                     colorkey = FALSE, labels=as.character(c(-4,-1.5,-0.8,0.8,1.5,4)),
                     margin=F, scales = list(draw = FALSE),main = list(label = "CAL_Age", cex = 1))

print(CAL2map, split = c(1, 2, 5, 2), more = TRUE)

## CAL Gender ##
CAL3 <- matrix(NA,168,n.iter)
for (i in 1:n.iter){
  CAL3[,i] <- CAL3init[(168*(i-1) + 1):(168*i),] 
}

CICAL3 <- matrix(NA,168,3)

for (i in 1:168){
  CICAL3[i,] <- quantile(CAL3[i,], c(0.025,0.5,0.975))
}


CAL3_d1 <- as.matrix(CICAL3[,2]); CAL3_d2 <- matrix(NA, 168, 1)
for (i in 1:168){
  if (CAL3_d1[i,] > 0){
    CAL3_d2[i,] <- CAL3_d1[i,] - CICAL3[i,1]
  }
  else {CAL3_d2[i,] <- CICAL3[i,3] - CAL3_d1[i,]}
}
CAL3map <- CAL3_d1/abs(CAL3_d2)
CAL3map <- matrix(CAL3map, t,s)

CAL3map <- levelplot(t(CAL3map),col.regions = Acol.regions,xlab="sites",ylab="teeth", at = c(-4,-1.5,-0.8,0.8,1.5,4),
                     colorkey = FALSE, labels=as.character(c(-4,-1.5,-0.8,0.8,1.5,4)),
                     margin=F, scales = list(draw = FALSE),main = list(label = "CAL_Gender", cex = 1))

print(CAL3map, split = c(2, 2, 5, 2), more = TRUE)

## CAL BMI ##
CAL4 <- matrix(NA,168,n.iter)
for (i in 1:n.iter){
  CAL4[,i] <- CAL4init[(168*(i-1) + 1):(168*i),] 
}

CICAL4 <- matrix(NA,168,3)

for (i in 1:168){
  CICAL4[i,] <- quantile(CAL4[i,], c(0.025,0.5,0.975))
}


CAL4_d1 <- as.matrix(CICAL4[,2]); CAL4_d2 <- matrix(NA, 168, 1)
for (i in 1:168){
  if (CAL4_d1[i,] > 0){
    CAL4_d2[i,] <- CAL4_d1[i,] - CICAL4[i,1]
  }
  else {CAL4_d2[i,] <- CICAL4[i,3] - CAL4_d1[i,]}
}
CAL4map <- CAL4_d1/abs(CAL4_d2)
CAL4map <- matrix(CAL4map, t,s)

CAL4map <- levelplot(t(CAL4map),col.regions = Acol.regions,xlab="sites",ylab="teeth", at = c(-4,-1.5,-0.8,0.8,1.5,4),
                     colorkey = FALSE, labels=as.character(c(-4,-1.5,-0.8,0.8,1.5,4)),
                     margin=F, scales = list(draw = FALSE),main = list(label = "CAL_BMI", cex = 1))
print(CAL4map, split = c(3, 2, 5, 2), more = TRUE)

## CAL Smoker ##
CAL5 <- matrix(NA,168,n.iter)
for (i in 1:n.iter){
  CAL5[,i] <- CAL5init[(168*(i-1) + 1):(168*i),] 
}

CICAL5 <- matrix(NA,168,3)

for (i in 1:168){
  CICAL5[i,] <- quantile(CAL5[i,], c(0.025,0.5,0.975))
}

CAL5_d1 <- as.matrix(CICAL5[,2]); CAL5_d2 <- matrix(NA, 168, 1)
for (i in 1:168){
  if (CAL5_d1[i,] > 0){
    CAL5_d2[i,] <- CAL5_d1[i,] - CICAL5[i,1]
  }
  else {CAL5_d2[i,] <- CICAL5[i,3] - CAL5_d1[i,]}
}
CAL5map <- CAL5_d1/abs(CAL5_d2)
CAL5map <- matrix(CAL5map, t,s)

CAL5map <- levelplot(t(CAL5map),col.regions = Acol.regions,xlab="sites",ylab="teeth", at = c(-4,-1.5,-0.8,0.8,1.5,4),
                     colorkey = FALSE, labels=as.character(c(-4,-1.5,-0.8,0.8,1.5,4)),
                     margin=F, scales = list(draw = FALSE),main = list(label = "CAL_Smoker", cex = 1))
print(CAL5map, split = c(4, 2, 5, 2), more = TRUE)

## CAL HbA1c ##
CAL6 <- matrix(NA,168,n.iter)
for (i in 1:n.iter){
  CAL6[,i] <- CAL6init[(168*(i-1) + 1):(168*i),] 
}

CICAL6 <- matrix(NA,168,3)

for (i in 1:168){
  CICAL6[i,] <- quantile(CAL6[i,], c(0.025,0.5,0.975))
}

CAL6_d1 <- as.matrix(CICAL5[,2]); CAL6_d2 <- matrix(NA, 168, 1)
for (i in 1:168){
  if (CAL6_d1[i,] > 0){
    CAL6_d2[i,] <- CAL6_d1[i,] - CICAL6[i,1]
  }
  else {CAL6_d2[i,] <- CICAL6[i,3] - CAL6_d1[i,]}
}
CAL6map <- CAL6_d1/abs(CAL6_d2)
CAL6map <- matrix(CAL6map, t,s)

CAL6map <- levelplot(t(CAL6map),col.regions = Acol.regions,xlab="sites",ylab="teeth", at = c(-4,-1.5,-0.8,0.8,1.5,4),
                     colorkey = FALSE, labels=as.character(c(-4,-1.5,-0.8,0.8,1.5,4)),
                     margin=F, scales = list(draw = FALSE),main = list(label = "CAL_HbA1c", cex = 1))
print(CAL6map, split = c(5, 2, 5, 2), more = TRUE)
```

Now, the following plots (Figure A.3 in the Supplementary materials) represent residuals (real values - fitted values) for PPD & CAL of randomly picked patient.

```{r, echo=FALSE}
setwd('C:/Users/inkku/Downloads')
load("resd_PPD.rda"); load("resd_CAL.rda")
```

```{r, fig.align='center', fig.width=6, fig.height=3}
library(lattice);library(viridisLite)
t = 28; s = 6; b = 2; n = 290;

PPD_resd <- array(resd_PPD[,38], dim = c(t,s))
CAL_resd <- array(resd_CAL[,38], dim = c(t,s))


col <- viridis(100)

PPD <- array(t(PPD_resd), dim = c(6,28,1))
dimnames(PPD) <- list(c(1:6),c(1:28),c("PPD residual"))
levelplot(PPD,par.strip.text=list(cex=0.55), xlab="Sites",ylab="Teeth", scales = list(draw = FALSE), col.regions = col)

CAL <- array(t(CAL_resd), dim = c(6,28,1))
dimnames(CAL) <- list(c(1:6),c(1:28),c("CAL residual"))
levelplot(CAL,par.strip.text=list(cex=0.55), xlab="Sites",ylab="Teeth", scales = list(draw = FALSE), col.regions = col)
```














